---
title: "UWES R Tutorial"
author: "Andrew Girgis"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

- Introduction to R
- loading and viewing data
- Data manipulation

## Download R and RStudio

R is the programming language and RStudio is an integrated development environment (IDE) for R. You need to install both before you are ready to start programming.

- **Download R:** [https://cran.r-project.org/](https://cran.r-project.org/)
- **Download RStudio:** [https://posit.co/download/rstudio-desktop/](https://posit.co/download/rstudio-desktop/)

## Getting started with R

R and RStudio are free and open source software environment for statistical computing and graphics. It compiles and runs on a wide variety of platforms, including Windows and MacOS.

## RStudio IDE

RStudio is an integrated development environment (IDE) for R. It includes a console, syntax-highlighting editor that supports direct code execution, as well as tools for plotting, history, debugging and workspace management.

## Why use R?

- R’s role in data science and economic research.
- Comparison with other tools like Excel, Python, and Stata.
- Real-world examples of economists and analysts using R.


## R Basics - Syntax & Data Types

- Variables and assignment 
```{r, echo=TRUE}
x <- 10
```
- Data types: Numeric, Character, Logical, Factor
```{r, echo=TRUE}
class(x)
```
- Printing values
```{r, echo=TRUE}
print(x)
```
- Comments in R 
```{r, echo=TRUE}
# This is a comment
```

## R Basics - Syntax & Data Types

- Arithmetic operations
```{r, echo=TRUE}
y <- 20
z <- x + y
print(z)
```

## Data Structures in R

- Vectors (c(1, 2, 3))
```{r}
v <- c(1, 2, 3)
```

- Matrices (matrix(1:9, nrow = 3))
```{r}
m <- matrix(1:9, nrow = 3)
```

- Data frames (data.frame(name=c("A", "B"), age=c(21, 25)))
```{r}
df <- data.frame(name=c("A", "B"), age=c(21, 25))
```

- Lists (list(name="John", age=30, salary=4000))
```{r}
l <- list(name="John", age=30, salary=4000)
```


## Importing & Manipulating Data

- Loading built-in datasets (data(mtcars))
```{r}
data(mtcars)
```
- Reading CSV files (read.csv("data.csv"))
- Subsetting data (df[1:10,])
- Filtering with logical conditions (subset(df, age > 25))

## Basic Data Analysis in R

- Summary statistics (summary(mtcars))
- Calculating mean, median, standard deviation (mean(mtcars$mpg))
- Simple plotting (hist(mtcars$mpg))

## Hands-on Exercise
Activity:
Provide a small dataset for participants to load into R.
Have them calculate basic statistics (mean, median, standard deviation).
Challenge: Plot a histogram or scatterplot.
